### 2021-02-12 Modules/sec_prlm_FIM.tex
Sent as draft No.2

In the parameter estimation problems, we obtain information about the parameter from asample of data coming from the underlying probability distribution. 
A natural question is:how much information can a sample of data provide about the unknown parameter? 
This section introduces such a measure for information, and we can also see that this informationmeasure can be used to find bounds on the variance of estimators, and it can be used toapproximate the sampling distribution of an estimator obtained from a large sample, andfurther be used to obtain an approximate confidence interval in case of large sample.
In this section, we consider a random variable $X$ for which the pdf or pmf is $f\rBrace{x|\theta}$, wehre $\theta$ is an unknown parameter and $\theta\in\Theta$, with $\Theta$ is the parameter space.
Intuitively, if an event has small probability, then the occurrence of this eventbrings us much information. 
For a random variable $X \sim f\rBrace{x|\theta}$, if $\theta$ were the true value ofthe parameter, the likelihood function should take a big value, or equivalently, the derivativelog-likelihood function should be close to zero, and this is the basic principle of maximumlikelihood estimation.
We define $l\rBrace{x|\theta} = \log{f\rBrace{x|\theta}}$ as the log-likelihood function, and
    l^{\prime}\rBrace{x|\theta} = \frac{\partial}{\partial\theta}\log{f\rBrace{x|\theta}}=\frac{f^{\prime}\rBrace{x|\theta}}{f\rBrace{x|\theta}}
where $f^{\prime}\rBrace{x|\theta}$ is the derivative of $f\rBrace{x|\theta}$ with respect to $\theta$.
Similarly, we denote the second order derivative of $f\rBrace{x|\theta}$ with respect to $\theta$ as $f^{\prime\prime}\rBrace{x|\theta}$.
According to the above analysis, if $l^{\prime}\rBrace{X|\theta}$ is close to zero, then it is expected, thus the random variable does not provide much information about $\theta$;
on the other hand, if $\abs{l^{\prime}\rBrace{X|\theta}}$ or $\abs{l^{\prime}\rBrace{X|\theta}}^{2}$ is large, the random variable provides much information about $\theta$.
Thus, we can use $\abs{l^{\prime}\rBrace{X|\theta}}^{2}$ to measure the amount of information provided by $X$.
however, since $X$ is a random variable, we should consider the average case. Thus, we introduce the following definition:
\begin{definition}
\emph{Fisher information} (for $\theta$) contained in the random variable $X$ is defined as:
\label{eq_prlm_FI_def}
    I\rBrace{\theta}=E_{\theta}\cBrace{\abs{l^{\prime}\rBrace{X|\theta}}^{2}}=\int{}\abs{l^{\prime}\rBrace{X|\theta}}^{2}f\rBrace{x|\theta}dx.
\end{definition}
We assume that we can exchange the order of differentiation and integration, then
    \int{}f^{\prime}\rBrace{x|\theta}dx=\frac{\partial}{\partial\theta}\int{}f\rBrace{x|\theta}dx=0.
Similarly,
    \int{}f^{\prime\prime}\rBrace{x|\theta}dx=\frac{\partial^{2}}{\partial\theta^{2}}\int{}f\rBrace{x|\theta}dx=0.
It is easy to see that
    E_{\theta}\cBrace{l^{\prime}\rBrace{X|\theta}}=\int{}l^{\prime}\rBrace{x|\theta}f\rBrace{x|\theta}dx=\int{}\frac{f^{\prime}\rBrace{x|\theta}}{f\rBrace{x|\theta}}f\rBrace{x|\theta}dx=\int{}f^{\prime}\rBrace{x|\theta}=0
Therefore, the definition of Fisher information \eqref{eq_prlm_FI_def} can be rewritten as
\label{eq_prlm_IeqVar}
    I\rBrace{\theta}=\text{Var}_{\theta}\cBrace{l^{\prime}\rBrace{X|\theta}}
Also, notice that
    l^{\prime\prime}\rBrace{x|\theta}=\frac{\partial}{\partial\theta}\vBrace{\frac{f^{\prime}\rBrace{x|\theta}}{f\rBrace{x|\theta}}}=\frac{f^{\prime\prime}\rBrace{x|\theta}f\rBrace{x|\theta}-\abs{f^{\prime}\rBrace{x|\theta}}^{2}}{\abs{f\rBrace{x|\theta}}^{2}}=\frac{\abs{f^{\prime\prime}\rBrace{x|\theta}}^{2}}{\abs{f\rBrace{x|\theta}}^{2}}-\abs{l^{\prime}\rBrace{x|\theta}}^{2}.
Therefore,
    E_{\theta}\cBrace{l^{\prime\prime}\rBrace{x|\theta}}=\int{\vBrace{\frac{\abs{f^{\prime\prime}\rBrace{x|\theta}}^{2}}{\abs{f\rBrace{x|\theta}}^{2}}-\abs{l^{\prime}\rBrace{x|\theta}}^{2}}f\rBrace{x|\theta}}=\int{f^{\prime\prime}\rBrace{x|\theta}dx}-E_{\theta}\cBrace{l^{\prime}\rBrace{X|\theta}}=-I\rBrace{\theta}.
Finally, we have another formula to calculate Fisher information:

@@

To summarize, we have three methods to calculate Fisher information: equations \eqref{eq_prlm_FI_def}, \eqref{eq_prlm_IeqVar} and \eqref{eq_prlm_Ifin}.
In any problems, using \eqref{eq_prlm_Ifin} is the most convenient choice.

@@

Suppose that we have a random sample $X_{1},\dots,X_{n}$ coming from a distribution for which the pdf or pmf is $f\rBrace{x|\theta},$ where the value of the parameter $\theta$ is unknown. 
We will show how to used Fisher information to determine the lower bound for the variance of an estimator of the parameter $\theta$.
Let $\hat{\theta}=r\rBrace{X_{1},\dots,X_{n}}=r\rBrace{\vecnot{X}}$ be an arbitrary estimator of $\theta$.
Assume $E_{\theta}\rBrace{\hat{\theta}}=m\rBrace{\theta}$ and the variance of $\hat{\theta}$ is finite.
Let us consider the random variable $l^{\prime}\rBrace{\vecnot{X}|\theta}$, it can be shown that $E_{\theta}\cBrace{l^{\prime}\rBrace{\vecnot{X}|\theta}}=0$ and that
\label{eq_prlm_covThetaLTag}
    \text{Cov}_{\theta}\cBrace{\hat{\theta},l^{\prime}\rBrace{\vecnot{X}|\theta}}=m^{\prime}\rBrace{\theta}.
By Cauchy-Schwartz inequality and the definition of $I_{n}\rBrace{\theta},$
    \vBrace{\text{Cov}_{\theta}\cBrace{\hat{\theta},l^{\prime}\rBrace{\vecnot{X}|\theta}}}^{2}\leq{}\text{Var}_{\theta}\cBrace{\hat{\theta}}\text{Var}_{\theta}\cBrace{l^{\prime}\rBrace{\vecnot{X}|\theta}}=\text{Var}_{\theta}\cBrace{\hat{\theta}}I_{n}\rBrace{\theta}
i.e.
    \vBrace{m^{\prime}\rBrace{\theta}}^{2}\leq{}\text{Var}_{\theta}\cBrace{\hat{\theta}}I_{n}\rBrace{\theta}=nI\rBrace{\theta}=\text{Var}_{\theta}{\cBrace{\hat{\theta}}}
Finally, we get the lower bound of variance of an arbitrary estimator $\hat{\theta}$ as
\label{eq_prlm_CramerRao_lim}
    \text{Var}_{\theta}{\cBrace{\hat{\theta}}}\geq{}\frac{\vBrace{m^{\prime}\rBrace{\theta}}^{2}}{nI\rBrace{\theta}}
The inequality \eqref{eq_prlm_CramerRao_lim} is called the \emph{information inequality}, and also known as the \emph{Cram\'er-Rao inequality} in honor of the Sweden statistician H. Cram\'er Indian statistician C. R. Rao who independently developed this inequality during the 1940s. 
The information inequality shows that as $I_{\theta}$ increases, the variance of the estimator decreases, therefore, the quality of the estimator increases, that is why the quantity is called “information”.
If $\hat{\theta}$ is an unbiased estimator, then $m\rBrace{\theta}=E_{\theta}\rBrace{\hat{\theta}}=\theta$, $m^{\prime}\rBrace{\theta}=1$.
Hence, by the information inequality, for unbiased estimator $\hat{\theta}$,

@@

    \test{Cov}_{\vecnot{\theta}}\cBrace{\hat{\vecnot{\theta}}\rBrace{\vecnot{X}}}\geq{}I^{-1}\rBrace{\vecnot{\theta}}

@@

>>>>

In the parameter estimation problems, we obtain information about the parameter from a sample of data coming from the underlying probability distribution. 
A natural question is: how much information can a sample of data provide about the unknown parameter? 
% This section introduces such a measure for information, and we can also see that this information measure can be used to find bounds on the variance of estimators, and it can be used to approximate the sampling distribution of an estimator obtained from a large sample, and further be used to obtain an approximate confidence interval in case of large sample.
To this end, the FIM, which is the common tools for assessing the amount of available information in the sampled data is overviewed in the following.
Consider a random variable $x$ for which the pdf is $f\rBrace{x|\theta}$, where $\theta$ is an unknown parameter and $\theta\in\Theta$, with $\Theta$ is the parameter space.
Intuitively, if an event has small probability, then the occurrence of this event brings us much information.
For a random variable $X \sim f\rBrace{x|\theta}$, if $\theta$ were the true value of the parameter, the likelihood function should take a big value, or equivalently, the derivative log-likelihood function should be close to zero, and this is the basic principle of maximum likelihood estimation.
We define $l\rBrace{x|\theta} = \log{f\rBrace{x|\theta}}$ as the log-likelihood function which follows that the Fisher information may be expressed \cite{van2004detection} as 
% We define $l\rBrace{x|\theta} = \log{f\rBrace{x|\theta}}$ as the log-likelihood function, and
% \begin{equation*}
%     l^{\prime}\rBrace{x|\theta} = \frac{\partial}{\partial\theta}\log{f\rBrace{x|\theta}}=\frac{f^{\prime}\rBrace{x|\theta}}{f\rBrace{x|\theta}}
% \end{equation*}
% where $f^{\prime}\rBrace{x|\theta}$ is the derivative of $f\rBrace{x|\theta}$ with respect to $\theta$.
% Similarly, we denote the second order derivative of $f\rBrace{x|\theta}$ with respect to $\theta$ as $f^{\prime\prime}\rBrace{x|\theta}$.
% According to the above analysis, if $l^{\prime}\rBrace{X|\theta}$ is close to zero, then it is expected, thus the random variable does not provide much information about $\theta$;
% on the other hand, if $\abs{l^{\prime}\rBrace{X|\theta}}$ or $\abs{l^{\prime}\rBrace{X|\theta}}^{2}$ is large, the random variable provides much information about $\theta$.
% Thus, we can use $\abs{l^{\prime}\rBrace{X|\theta}}^{2}$ to measure the amount of information provided by $X$.
% however, since $X$ is a random variable, we should consider the average case. Thus, we introduce the following definition:
% \begin{definition}
% \emph{Fisher information} (for $\theta$) contained in the random variable $X$ is defined as:
% \begin{equation}
% \label{eq_prlm_FI_def}
%     I\rBrace{\theta}=E_{\theta}\cBrace{\abs{l^{\prime}\rBrace{X|\theta}}^{2}}=\int{}\abs{l^{\prime}\rBrace{X|\theta}}^{2}f\rBrace{x|\theta}dx.
% \end{equation}
% \end{definition}
% We assume that we can exchange the order of differentiation and integration, then
% \begin{equation*}
%     \int{}f^{\prime}\rBrace{x|\theta}dx=\frac{\partial}{\partial\theta}\int{}f\rBrace{x|\theta}dx=0.
% \end{equation*}
% Similarly,
% \begin{equation*}
%     \int{}f^{\prime\prime}\rBrace{x|\theta}dx=\frac{\partial^{2}}{\partial\theta^{2}}\int{}f\rBrace{x|\theta}dx=0.
% \end{equation*}
% It is easy to see that
% \begin{equation*}
%     E_{\theta}\cBrace{l^{\prime}\rBrace{X|\theta}}=\int{}l^{\prime}\rBrace{x|\theta}f\rBrace{x|\theta}dx=\int{}\frac{f^{\prime}\rBrace{x|\theta}}{f\rBrace{x|\theta}}f\rBrace{x|\theta}dx=\int{}f^{\prime}\rBrace{x|\theta}=0
% \end{equation*}
% Therefore, the definition of Fisher information \eqref{eq_prlm_FI_def} can be rewritten as
% \begin{equation}
% \label{eq_prlm_IeqVar}
%     I\rBrace{\theta}=\text{Var}_{\theta}\cBrace{l^{\prime}\rBrace{X|\theta}}
% \end{equation}.
% Also, notice that
% \begin{equation*}
%     l^{\prime\prime}\rBrace{x|\theta}=\frac{\partial}{\partial\theta}\vBrace{\frac{f^{\prime}\rBrace{x|\theta}}{f\rBrace{x|\theta}}}=\frac{f^{\prime\prime}\rBrace{x|\theta}f\rBrace{x|\theta}-\abs{f^{\prime}\rBrace{x|\theta}}^{2}}{\abs{f\rBrace{x|\theta}}^{2}}=\frac{\abs{f^{\prime\prime}\rBrace{x|\theta}}^{2}}{\abs{f\rBrace{x|\theta}}^{2}}-\abs{l^{\prime}\rBrace{x|\theta}}^{2}.
% \end{equation*}
% Therefore,
% \begin{equation*}
%     E_{\theta}\cBrace{l^{\prime\prime}\rBrace{x|\theta}}=\int{\vBrace{\frac{\abs{f^{\prime\prime}\rBrace{x|\theta}}^{2}}{\abs{f\rBrace{x|\theta}}^{2}}-\abs{l^{\prime}\rBrace{x|\theta}}^{2}}f\rBrace{x|\theta}}=\int{f^{\prime\prime}\rBrace{x|\theta}dx}-E_{\theta}\cBrace{l^{\prime}\rBrace{X|\theta}}=-I\rBrace{\theta}.
% \end{equation*}
% Finally, we have another formula to calculate Fisher information:

@@

% To summarize, we have three methods to calculate Fisher information: equations \eqref{eq_prlm_FI_def}, \eqref{eq_prlm_IeqVar} and \eqref{eq_prlm_Ifin}.
% In any problems, using \eqref{eq_prlm_Ifin} is the most convenient choice.

@@

% Suppose that we have a random sample $X_{1},\dots,X_{n}$ coming from a distribution for which the pdf or pmf is $f\rBrace{x|\theta},$ where the value of the parameter $\theta$ is unknown. 
% We will show how to use Fisher information to determine the lower bound for the variance of an estimator of the parameter $\theta$.
% \par
% Let $\hat{\theta}=r\rBrace{X_{1},\dots,X_{n}}=r\rBrace{\vecnot{X}}$ be an arbitrary estimator of $\theta$.
% Assume $E_{\theta}\rBrace{\hat{\theta}}=m\rBrace{\theta}$ and the variance of $\hat{\theta}$ is finite.
% Let us consider the random variable $l^{\prime}\rBrace{\vecnot{X}|\theta}$, it can be shown that $E_{\theta}\cBrace{l^{\prime}\rBrace{\vecnot{X}|\theta}}=0$ and that
% \begin{equation}
% \label{eq_prlm_covThetaLTag}
%     \text{Cov}_{\theta}\cBrace{\hat{\theta},l^{\prime}\rBrace{\vecnot{X}|\theta}}=m^{\prime}\rBrace{\theta}.
% \end{equation}.
% By Cauchy-Schwartz inequality and the definition of $I_{n}\rBrace{\theta},$
% \begin{equation*}
%     \vBrace{\text{Cov}_{\theta}\cBrace{\hat{\theta},l^{\prime}\rBrace{\vecnot{X}|\theta}}}^{2}\leq{}\text{Var}_{\theta}\cBrace{\hat{\theta}}\text{Var}_{\theta}\cBrace{l^{\prime}\rBrace{\vecnot{X}|\theta}}=\text{Var}_{\theta}\cBrace{\hat{\theta}}I_{n}\rBrace{\theta}
% \end{equation*}
% i.e.
In practical applications, where signal is noised and parameters cannot be accurately estimated, having a closed form expression for the FIM naturally leads to seek for a performance bound that will help the user to know when available information is fully utilized.
% \begin{equation*}
%     \vBrace{m^{\prime}\rBrace{\theta}}^{2}\leq{}\text{Var}_{\theta}\cBrace{\hat{\theta}}I_{n}\rBrace{\theta}=nI\rBrace{\theta}=\text{Var}_{\theta}{\cBrace{\hat{\theta}}}
% \end{equation*}
% Finally, we get the lower bound of variance of an arbitrary estimator $\hat{\theta}$ as
Such a bound for unbiased estimators is \cite{van2004detection}
% \begin{equation}
% \label{eq_prlm_CramerRao_lim}
%     \text{Var}_{\theta}{\cBrace{\hat{\theta}}}\geq{}\frac{\vBrace{m^{\prime}\rBrace{\theta}}^{2}}{nI\rBrace{\theta}}
% \end{equation}
% where 
% The inequality \eqref{eq_prlm_CramerRao_lim} is called the \emph{information inequality}, and also known as the \emph{Cram\'er-Rao inequality} in honor of the Sweden statistician H. Cram\'er Indian statistician C. R. Rao who independently developed this inequality during the 1940s. 
% The information inequality shows that as $I_{\theta}$ increases, the variance of the estimator decreases, therefore, the quality of the estimator increases, that is why the quantity is called “information”.
% \par
% If $\hat{\theta}$ is an unbiased estimator, then $m\rBrace{\theta}=E_{\theta}\rBrace{\hat{\theta}}=\theta$, $m^{\prime}\rBrace{\theta}=1$.
% Hence, by the information inequality, for unbiased estimator $\hat{\theta}$,

@@

    \text{Cov}_{\vecnot{\theta}}\cBrace{\hat{\vecnot{\theta}}\rBrace{\vecnot{X}}}\geq{}I^{-1}\rBrace{\vecnot{\theta}}

@@

\subsection{Frequency Domain Cram\'er-Rao Bound for Gaussian Processes}
The interesting work of \cite{zeira1990frequency} (which also specifically states its contribution to localization applications) enables the generalization of the CRLB to the frequency domain, which is of high relevance to this work (see Sec.\ref{sec_FIM}).
Assuming Gaussian processes, the results of \cite{whittle1953analysis} are utilized together with further development to find that 
\begin{equation}\label{eq_FIM_kl_full_REAL}
    %\resizebox{.9\linewidth}{!}
    {
        \begin{split}
            J_{k,l}\rBrace{\vEta} 
            =&
            %\Re\cBrace{
            \frac{1}{2\pi}
            \int_{-\omega_{s}/2}^{\omega_{s}/2}
            {
            \frac{1}{\Phi\rBrace{\omega}}
            \mathfrak{F}^{*}\left\{
            \frac{\partial z(t)}{\partial\eta_{k}}
            \right\}
            \mathfrak{F}\left\{
            \frac{\partial z(t)}{\partial\eta_{l}}
            \right\}
            d\omega
            }%}
            \\ &+
            \frac{T}{4\pi}
            \int_{-\omega_{s}/2}^{\omega_{s}/2}
            \frac{1}{\Phi^{2}\rBrace{\omega}}
            \rBrace{\frac{\partial\Phi\rBrace{\omega}}{\partial\eta_{k}}}^{\ast}
            \frac{\partial\Phi\rBrace{\omega}}{\partial\eta_{l}}
            d\omega
        \end{split}
    }
which is conveniently generalized to the complex data case used in this work as \eqref{eq_FIM_kl_full}.

### 2021-02-12 Modules/sec_prlm_FIR_IIR.tex
Sent as draft No.2

A basic design entity in every signal processing scheme is the digital filter, which is the modern evolution of analog filters.
The design of the digital filter is a very established and thoroughly studied research area.
As stated in \cite{oppenheim1975digital} (and numerous other literature), two key filter types are used - i.e. Finite Impulse Response (FIR) and Infinite Impulse Response (IIR) filters.
This section will briefly overview the design methods and main flavours of each filter type and discuss in general of the pros and cons of them in order to provide proper motivation for the quest of finding the IIR analogy of the spatial FIR, as implied in Ch.~\ref{chap:intro}.
\subsection{FIR design}
The FIR filter is a merely a delay-and-sum mechanism, therefore its memory is finite and equal to the number of taps in the final filter's configuration.
Determining it coefficients is generally done using windowing (weighted element-wise multiplication and truncation) of desired impulse response which is derived by the inverse Fourier transform of the desired frequency response.
Many window types have been suggested in literature, where Bartlet, Hanning, Humming and Blackman are among the most basic ones, each with its pros and cons.
I a nutshell, using a rectangular window, achieves the narrowest main lobe but also the highest side-lobes.
The other window types, achieve lower side-lobes at the expense of wider main-lobe.
The user may also increase the filter's order if side-lobe attenuation is of high importance, thus achieving narrower main-lobe. 
\subsection{IIR design}
The infinite response of the IIR filters originates in the inherent feedback which reuses output in the computing part, generating a loop which theoretically takes all input past samples into account for each output sample.
As there are many flavours of the IIR filters, we choose to elaborate on the elliptic filter, being the more common candidate for comparison to the FIR in the literature.
The basic concept behind the elliptic filter design is the allowing of equi-ripple around both pass and stop band which both relaxes the filter's constrains comparing to Chebyshev and Butterworth methods and allows the narowest transition band between the three filters.
The designer sets the allowed ripple magnitudes in both pass and stop bands and positions the transition band according to his needs.
A closed for transformation, described in \cite{oppenheim1975digital} is then used to convert the prototype low-pass filter to band-pass, band-stop or high-pass filters.
\subsection{Comparison}
The main advantage of FIR filters, commonly required in signal processing systems is the obtainable linear phase feature. A well known \cite{oppenheim1975digital} fact in digital signal processing is that the filter latency may be frequency dependant, namely linearly dependent in the phase response derivative of the specific frequency.
Therefore, linear phase implies that the phase response's derivative (i.e. signal latency) is frequency independents, which prevent signal distortion.
The main advantage of IIR filters is that for a given set of frequency response specifications, the IIR filter will (in most cases) require substantially less coefficients, which may even reach few order of magnitude \cite{rabiner1974some}.
In this work, the signals are assumed to be narrow-band, which reduces the necessity of linear phase response, while when considering sensor arrays, the costs and the physical aperture are of very high importance, motivating the designers to try and reduce the number of elements in the array while still meeting the specifications.

>>>>

A basic design entity in every signal processing scheme is the digital filter, which is the modern evolution of analog filters.
The design of the digital filter is very well established and thoroughly studied research field.
As stated in \cite{oppenheim1975digital} (and numerous other literature), two key filter types are used - i.e. Finite Impulse Response (FIR) and Infinite Impulse Response (IIR) filters.
This section will briefly overview the design methods and the analogy of ULA based beamforming to the temporal FIR filtering, as implied in Ch.~\ref{chap:intro}.
\par Each digital filters belong in one of two architectures
\begin{itemize}
    \item \textbf{FIR design}\\
    The FIR filter is a merely a delay-and-sum mechanism, therefore its memory is finite and equal to the number of taps in the final filter's configuration as in Fig.~\ref{fig_FIR_arch}.
    \item \textbf{IIR design}\\
    IIR filters may be viewed as a generalization of the FIR delay-and-sum architecture, where delayed instances of the array's output are fed back to the summation as in Fig.~\ref{fig_IIR_arch}, theoretically generating an infinite loop where each output sample is affected by all past input samples. 
\end{itemize}
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.7\linewidth, 
        %grid, 
        tics=10,trim=0 0 0 0]{./Media/BASIC_FIR_FILTER_ARCH.png}
            \put (13, 36){\footnotesize{$s$}}
            \put (22, 22){\footnotesize{$\beta_{0}$}}
            \put (38.75, 22){\footnotesize{$\beta_{1}$}}
            \put (55.5, 22){\footnotesize{$\beta_{2}$}}
            \put (80.75, 22){\footnotesize{$\beta_{N-1}$}}
            \put (85, 10){\footnotesize{$z$}}
        \end{overpic}
    \end{center}
    \caption{An FIR based filtering of signal $s$ with the $\vecnot{\beta}$ as the coefficients, generating a filtered signal $z$.}
    \label{fig_FIR_arch}
\end{figure}
\begin{figure}[h!]
    \begin{center}
        \begin{overpic}[width=0.7\linewidth, 
        % grid, 
        tics=10,trim=0 0 0 0]{./Media/BASIC_IIR_FILTER_ARCH.png}
            \put (60, 50){\footnotesize{$\beta_{0}$}}
            \put (60, 30){\footnotesize{$\beta_{1}$}}
            \put (60, 10){\footnotesize{$\beta_{2}$}}
            \put (36, 30){\footnotesize{$\alpha_{1}$}}
            \put (36, 10){\footnotesize{$\alpha_{2}$}}
            \put (10, 50){\footnotesize{$s$}}
            \put (85, 50){\footnotesize{$z$}}
            \put (47.5, 34.5){\footnotesize{$z^{-1}$}}
            \put (47.5, 14.5){\footnotesize{$z^{-1}$}}
        \end{overpic}
    \end{center}
    \caption{Direct form II $2^{nd}$ order IIR architecture, where the FIR part is the $\vecnot{\beta}$ coefficients and the recursive part is implemented via the $\vecnot{\alpha}$ set.}
    \label{fig_IIR_arch}
\end{figure}
\par In the following, we revisit the basic interpretation of the ULA as the spatial equivalent to the temporal FIR filter \cite{van1988beamforming} where DOA is shown to be the matching spatial entity to the temporal frequency, hence the \emph{conventional beamformer} \cite{van2004optimum} is merely the spatial version of temporal domain FIR filtering. To this end, a brief overview on ULA beamforming and its resemblance to FIR filter design are presented next.
\par Consider an $N$-element ULA with inter-element spacing $d$, where its $n$'th sensor is positioned at $\vpi{n},\;\text{for}\; n=0,\ldots,N-1$. We set $\vpi{0}$ as the axis reference point and assume that the target of interest is positioned at $\vpt$.
Focusing on a far field localization problem, DOA and target range are to be estimated. Without loss of generality, aiming to simplify the exposition, we assume $2\text{D}$ planar problem, where DOA is described by a single angle $\thetaD$, measured from the array's broadside.
For simplicity, we assume an anechoic environment, an array of identical omni-directional sensors and a stationary target of interest.
Inspired by radar based applications, we also place a transmitter at $\vpi{0}$, assuming the transmitted signal $s$, is reflected back from the target and re-impinges the array, with a total time delay of $\tau_{\text{pd}}=2R/c$ seconds, where $R = \norm{\vpt-\vpi{0}}$ is the target's range and $c$ represents the propagation velocity of the signal in the medium.
\par Let $x_{n}(t)$ be the measured signal at the $n$'th sensor
\begin{equation}
x_{n}(t) = gs\Brack{t-\tau_{pd}-\tau_{n}},
\label{eqn:noFeedbackULA_singleSensor_temporal}
\end{equation}
where $\tau_{n}=nd\cos\Brack{\thetaD}/c$ represents the time difference of arrival between the $n$'th sensor and the reference sensor and $g$, being a scalar in an anechoic environment, is the channel's gain, related to both propagation and the target's radar cross section (RCS).
Defining $\vecnot{x}\rBrace{t}\triangleq\vBrace{x_{0}\rBrace{t}\hdots{}x_{N-1}\rBrace{t}}^{T}$ and its Fourier transform, $\vecnot{\F{x}}\rBrace{\omega}\triangleq\vBrace{X_{0}\rBrace{\omega},\hdots,X_{N-1}\rBrace{\omega}}^{T}$, one may write 
\[
\vecnot{\F{x}}\rBrace{\omega,\theta_{d}}=g\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}\rBrace{\theta}}}
\]
where $S\rBrace{\omega}$ is the Fourier transform of $s\rBrace{t}$ and $\vdO$ denotes the steering vector whose $n$'th element is
\begin{equation}
    \label{eq:d}
    d_{n}\rBrace{\omega,\theta_{d}} = \exp{\rBrace{-j\omega\tau_{n}\rBrace{\theta}}}.
\end{equation}
Denoting the beamformer's weights as $\vBeta\rBrace{\omega}$ and the beamformer's output as $z$, we express the latter in the frequency domain
\begin{equation}
    \label{eq_Z}
    Z\rBrace{\omega,\theta_{d}} = g\vecnot{\beta}^{H}\rBrace{\omega}\vecnot{d}\rBrace{\omega}\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}}}.
\end{equation}
Defining the electric phase to be
\begin{equation}\label{eq:thetaULA}
\theta=\omega{d\cos\Brack{\thetaD}}/{c},
\end{equation}
we rewrite \eqref{eq_Z} as 
\[
Z\rBrace{\omega,\theta} = g\F{s}\rBrace{\omega}\exp{\rBrace{-j\omega\tau_{pd}}}\sum_{n=0}^{N-1}\beta^{*}_{n}\rBrace{\omega}\exp\Brack{-jn\theta},
\]
hence in the ULA case, aiming for a desired spatial response, the weights vector $\vecnot{\beta}\rBrace{\omega}$ configuration is mathematically equivalent to an FIR filter design~\cite{van1988beamforming,benesty2018}. 
Assuming narrowband stimuli signals, we suppress $\omega$ dependency in the notation throughout the rest of this paper, where possible.
\par It is well known \cite{rabiner1974some} that IIR filters have several appealing advantages over their matching (in number of elements) FIR versions.
A well known \cite{oppenheim1975digital} fact in digital signal processing is that the filter latency may be frequency dependant, namely linearly dependent in the phase response derivative of the specific frequency.
Therefore, linear phase implies that the phase response's derivative (i.e. signal latency) is constant which implies all input frequencies will reside in the processor for the exact same time, preventing dispersion which practically causes pulse widening.
Due to this important property, many applications which are sensitive to signal distortions, especially communication based systems, chose the FIR architecture.
Acquainted with the analogy between frequency and DOA, it seems that when localization applications are considered, where the DOA resolution criteria is more important than the consistency of detection latency between separate DOAs, the linear phase property is less relevant.
\par The main advantage of IIR filters is that for a given set of frequency response specifications, the IIR filter will (in most cases) require substantially less coefficients, which may even reach few order of magnitude \cite{rabiner1974some}.
Concluding that spatial linear phase response in DOA domain is not a necessity, combined with the known efficiency of the IIR architecture and the need for smaller and cheaper arrays, the motivation for finding the spatial version of the temporal IIR is complete.
\par In standard radar signal processing schemes, a waveform is transmitted to, and reflected from the target of interest. Then, the reflected signal is processed by the radar reception array in order to estimate the target's dynamics (e.g., DOA, range, velocity etc.).
As opposed to the standard scheme, we suggest a continuous re-transmission of the signal and its echoes back to the platform, generating a spatial feedback loop between the array and the target.
Another deviation from traditional radar processing, used to simplify the exposition, is using continuous-wave (CW) stimuli, rather than using pulse based signals.
